---
title: "Benchmarking TreeMMM Against Regression Baselines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking TreeMMM Against Regression Baselines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = requireNamespace("lightgbm", quietly = TRUE) &&
             requireNamespace("lme4",     quietly = TRUE)
)
```

# Benchmarking TreeMMM head to head

The Python TreeMMM paper benchmarks the tree-based pipeline against three
regression families on four synthetic DGPs. This vignette runs the same
benchmark on the pharma DGP at a small scale so the result fits in a
local R session.

## Setup

```{r}
library(treemmm)

ds <- generate_pharma_dataset(
  n_customers  = 100,
  n_periods    = 12,
  random_state = 42
)

cfg <- run_config(
  columns = column_spec(
    customer_id  = ds$columns$customer_id,
    time_col     = ds$columns$time_col,
    outcome_col  = ds$columns$outcome_col,
    promo_vars   = ds$columns$promo_vars,
    control_vars = ds$columns$control_vars
  ),
  objective    = "auto",
  random_state = 42L
)
```

## TreeMMM

```{r}
tree_result <- treemmm_run(ds$df, cfg)
tree_shares <- unlist(tree_result$attribution_shares)
```

## GLMM-Naive (main effects only, random customer intercept)

```{r}
glmm_naive <- fit_glmm_naive(ds$df, cfg)
naive_shares <- unlist(glmm_naive$attribution_shares)
```

## GLMM-Oracle (planted interactions given)

```{r}
glmm_oracle <- fit_glmm_oracle(ds$df, cfg, ds$ground_truth$interactions)
oracle_shares <- unlist(glmm_oracle$attribution_shares)
```

## Distributional-GLM ablation

This isolates "did the wrong link function fix the problem?". A
distributional GLM with the matched Poisson likelihood replaces the
log1p transform used by GLMM-Naive.

```{r}
glmm_dist <- fit_glmm_distributional(ds$df, cfg)
dist_shares <- unlist(glmm_dist$attribution_shares)
```

## Compare attribution shares

Each baseline's recovered share per promotional channel, side by side
with the planted DGP reference share:

```{r}
promo <- ds$columns$promo_vars
true_shares <- unlist(ds$ground_truth$attribution_shares)[promo]

cmp <- data.frame(
  channel    = promo,
  truth      = round(true_shares,                   3),
  treemmm    = round(tree_shares[promo],            3),
  glmm_naive = round(naive_shares[promo],           3),
  glmm_orcl  = round(oracle_shares[promo],          3),
  glmm_dist  = round(dist_shares[promo],            3),
  row.names  = NULL
)
cmp
```

## Attribution-share MAPE (lower is better)

```{r}
mape <- function(recovered, truth) {
  active <- truth > 0.005
  mean(abs(recovered[active] - truth[active]) / truth[active])
}

c(
  treemmm    = mape(tree_shares[promo],   true_shares),
  glmm_naive = mape(naive_shares[promo],  true_shares),
  glmm_orcl  = mape(oracle_shares[promo], true_shares),
  glmm_dist  = mape(dist_shares[promo],   true_shares)
)
```

At small scale results are noisy. The Python paper's headline values
(3,000 customers x 36 months, N=5 seeds) are TreeMMM 17.9 +/- 0.2 vs.
GLMM-Naive 22.2 +/- 0.3.

## Marginal ROI

```{r}
mroi_ranking(tree_result, channels = promo)
```

Compare the model's mROI ranking against the DGP's:

```{r}
bench <- mroi_benchmark(tree_result, ds)
bench$rank_correlation
bench$direction_accuracy
bench$table
```
