---
title: "Playing with the DGP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Playing with the DGP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = TRUE
)
```

# Inspecting and tuning the synthetic DGPs

TreeMMM ships with six synthetic DGPs, each parameterizable along the same
axes: number of customers, number of periods, random seed, and (for some)
adstock decay. This vignette shows how to inspect what each DGP plants and
how to tune the knobs.

## Available DGPs

```{r}
library(treemmm)

# Four headline DGPs
ds_pharma <- generate_pharma_dataset(n_customers = 30, n_periods = 12)
ds_cpg    <- generate_cpg_dataset(n_customers = 30, n_periods = 12)
ds_saas   <- generate_saas_dataset(n_customers = 30, n_periods = 12)
ds_lin    <- generate_linear_dataset(n_customers = 30, n_periods = 12)

# Two specialty DGPs
ds_adstock <- generate_pharma_adstock_dataset(n_customers = 30, n_periods = 12)
ds_geo     <- generate_geo_panel_dataset(n_regions = 30, n_weeks = 12)

# Each returns the same structure
lapply(list(ds_pharma, ds_cpg, ds_saas, ds_lin),
       function(d) c(rows = nrow(d$df), channels = length(d$columns$promo_vars)))
```

## Inspecting the planted truth

`ground_truth$attribution_shares` is the reference each model is benchmarked
against:

```{r}
sort(unlist(ds_pharma$ground_truth$attribution_shares), decreasing = TRUE)
```

`ground_truth$customer_sensitivities` gives the per-customer per-channel
sensitivity vector drawn from each segment's multivariate normal:

```{r}
str(head(ds_pharma$ground_truth$customer_sensitivities, 2))
```

`ground_truth$interactions` lists the planted channel interactions:

```{r}
ds_pharma$ground_truth$interactions
```

## Heterogeneous customer sensitivity (HCS)

The pharma DGP plants a rheumatology/dermatology split. The mean
sensitivity vectors per segment determine the spread:

```{r}
sens <- ds_pharma$ground_truth$customer_sensitivities

# Rheum vs derm rep_visits sensitivity
rheum_cids <- ds_pharma$df[specialty == "rheumatology", unique(customer_id)]
derm_cids  <- ds_pharma$df[specialty == "dermatology",  unique(customer_id)]

rheum_rep <- vapply(rheum_cids, function(id) sens[[id]][["rep_visits"]],
                    numeric(1))
derm_rep  <- vapply(derm_cids,  function(id) sens[[id]][["rep_visits"]],
                    numeric(1))

c(rheum_mean = mean(rheum_rep), derm_mean = mean(derm_rep))
```

The DGP plants rheum mean 1.6 / derm mean 0.4 for rep_visits, so at any
nontrivial sample size the segment means should differ by roughly that
amount.

## Tuning sample size

The default sizes are small for fast tests. Headline-scale defaults match
the Python paper benchmark (3,000 x 36):

```{r, eval = FALSE}
# Headline scale (slow!  this is the actual paper scale)
ds_full <- generate_pharma_dataset(
  n_customers  = 3000,
  n_periods    = 36,
  random_state = 42
)
```

## Tuning the random seed

Two runs at the same seed produce identical panels:

```{r}
a <- generate_pharma_dataset(n_customers = 20, n_periods = 6, random_state = 7)
b <- generate_pharma_dataset(n_customers = 20, n_periods = 6, random_state = 7)
identical(a$df, b$df)
```

Different seeds produce different panels but the planted attribution
shares converge to the same values at scale (Monte Carlo error
diminishes with `n_customers x n_periods`).

## Adstock variant

The pharma-with-adstock DGP plants geometric carryover on each channel.
The raw and adstocked values are both available:

```{r}
ds_adstock$columns$adstock_decays
head(ds_adstock$df[, .(customer_id, period, rep_visits, rep_visits_raw)])
```

The model sees `rep_visits` (the cumulative-exposure column); the analyst
can inspect `rep_visits_raw` to verify the carryover was planted as
expected.

## Reproducibility caveat across R and Python

R's Mersenne Twister and Python's PCG64 produce different samples at the
same numeric seed. The DGP coefficients and structural form match the
Python implementation exactly; the planted reference attribution shares
converge to the Python ones within Monte Carlo error (< 0.5 pp at
headline scale). See `SPEC.md` in the repo for the formal specification
both implementations target.
