---
title: "TreeMMM Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TreeMMM Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = requireNamespace("lightgbm", quietly = TRUE)
)
```

# A five-minute tour of TreeMMM

TreeMMM is a tree-based Marketing Mix Modeling tool. Given a panel of
customer-by-period observations and a set of promotional columns, it
fits a gradient-boosted tree model, computes SHAP-derived attribution
shares per promotional channel, and produces marginal-ROI response curves.

This vignette walks through the smallest end-to-end run.

## Generate a synthetic panel

The package ships with six synthetic data-generating processes (DGPs)
with known ground truth. We use the pharma DGP at a small scale:

```{r}
library(treemmm)

ds <- generate_pharma_dataset(
  n_customers  = 50,
  n_periods    = 12,
  random_state = 42
)

# The result is a list with three elements:
names(ds)
```

`ds$df` is the panel; `ds$columns` is the role mapping; `ds$ground_truth`
contains the planted attribution shares (the answer the model is trying
to recover):

```{r}
head(ds$df)
ds$columns
ds$ground_truth$attribution_shares
```

## Configure the pipeline

`run_config()` packages the pipeline parameters; `column_spec()` describes
which column plays which role. The defaults pick a sensible objective
based on the outcome distribution.

```{r}
cfg <- run_config(
  columns = column_spec(
    customer_id  = ds$columns$customer_id,
    time_col     = ds$columns$time_col,
    outcome_col  = ds$columns$outcome_col,
    promo_vars   = ds$columns$promo_vars,
    control_vars = ds$columns$control_vars
  ),
  objective       = "auto",
  n_optuna_trials = 2L,
  n_folds         = 3L,
  min_train_frac  = 0.5,
  random_state    = 42L
)
```

## Run

```{r}
result <- treemmm_run(ds$df, cfg)
```

## Inspect the result

The recovered attribution shares per channel:

```{r}
sort(unlist(result$attribution_shares), decreasing = TRUE)
```

Per-fold performance metrics:

```{r}
do.call(rbind, lapply(result$fold_metrics, as.data.frame))
```

## What next

* See `vignette("benchmark")` for a head-to-head comparison against
  the GLMM-Naive and GLMM-Oracle baselines.
* See `vignette("dgp_play")` for how to tune DGP parameters and
  inspect the planted ground truth.
* `simulate_response()` and `mroi_ranking()` show how to read marginal
  ROI off the fitted model.
